Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-11-30 inst/paper/main.tex
minor

\documentclass[USenglish,cleveref, autoref, thm-restate]{lipics-v2019}
\usepackage{listings,hyperref,multirow,paralist,xspace,url}
\input{macro}

@@

\medskip\noindent{\it Eval in R.} The reflective interface exposed by
R is somewhat complex. The core library exports four functions with
slightly different semantics, \c{eval}, \c{evalq}, \c{eval.parent},
and \c{local}. \Eval is the more general function, and, unlike in
JavaScript, it takes three arguments, an expression, its environment
and the enclosing environment. The following discussion simplify some
details which are not relevant to this paper. For the interested
reader we recommend the excellent book by Hadley
Wickham~\cite{hadley}. The definition of \eval starts as follows:

@@

\medskip\noindent\emph{CRAN Packages.} The packages that are included
in this study are the top \CorpusPackages packages based on their
reverse dependencies. The set of reverse dependencies for some package
\emph{P} is transitive closure of packages that import \emph{P}. The
hypothesis underlying that choice is that packages with a large set of
reverse depencies are likely to be higher in quality and their tests
will provide better coverage. The packages are implemented in R with
some C and Fortran code. Using \c{cloc}, we measure a total
\CorpusRCodeRnd lines of R code and \CorpusNativeCodeRnd lines of
native code. For each package, we use its runnable code (tests, etc.)
to compute code coverage. On average, the cde coverage is
\CorpusMeanExprCoverage which is acceptable without being exhaustive.
Figure~\ref{fig:corpus} shows these packages, the size of the dots
reflects the project's size in lines of code. The x-axis indicates
code coverage in percents and the y-axis gives the number of call
sites to \eval that were traced, in log scale. Dotted lines indicate
means. Packages with over \CorpusEvalsPackageTreshold eval call sites
are named.

@@

\medskip\noindent\emph{Kaggle Scripts.} Kaggle is an online platform
for data-science and machine-learning. The website allows people to
submit data-analysis problems, users compete to find the best
solution. The solutions are uploaded to the platform as either plain
scripts or notebooks. We chose one of the most popular competition,
predicting the survival of passengers the
Titanic\footnote{\url{https://www.kaggle.com/c/titanic}}. Unlike CRAN,
Kaggle is not curated. After downloading the \CorpusKaggle solutions
and extracting the R code, we found that \CorpusDuplicatedKaggle were
duplicates. From the remaining \CorpusRunnableKaggle solutions,
\CorpusFailedKaggle failed to execute. Next to various runtime
exceptions, common problems were parse errors and misspelled package
names. The final set contains \CorpusFinishedKaggle programs
implemented in \CorpusFinishedKaggleCodeRnd lines of R code.

@@

\begin{figure}[!h]\hspace{-5mm}\includegraphics[width=1.05\linewidth]
{pipeline.pdf}\caption{Analysis Pipeline}\label{fig:pipeline}
\end{figure}

@@

\medskip\noindent\emph{Limitations.} Dynamic analysis can only observe
calls that are triggered by the program's input. We believe that
focusing on R packages with high code coverage does mitigate this to
some extent. The results we report here were obtained with R's
bytecode compiler turned off, this should not affect the results as
the compiler does not optimize \eval.

@@

\newcommand{\mypara}[1]{\medskip\noindent\emph{#1}\xspace}

@@

\newcommand{\NOTE}[1]{{\it Note: #1}\xspace}

@@

\subsubsection{Unnecessary use of eval}
Similarly to JavaScript, there are also unnecessary uses of \eval. For
example, the \c{PerformanceAnalytics} package contains a function
\c{chart.QQPlot} that uses \eval to resolve a string into function and
another to call it and assign its results into a variable:
\begin{lstlisting}
function (R, d="norm", dp, ...) {
q.f <- eval(parse(text=paste("q",d,sep="")))
z <- NULL
eval(parse(text=paste("z<-q.f(",dp,",...)")))
}
\end{lstlisting}
  In both cases, there is no need for \eval:
\begin{lstlisting}
function (R, d="norm", dp, ...) {
q.f <- get(paste0("q",d))
z <- q.f(dp, ...)
}
\end{lstlisting}
  or even to a oneliner \c{do.call(paste0("q",d), as.list(dp, ...))}.

@@

@@

@@

\medskip\noindent\emph{Discussion:} Comparing the use of \eval in R
and JavaScript is instructive. The use of \eval is more widespread in
data science code than it was in JavaScript. Any run of an R program
seems to trigger \eval; moreover package programmers use \eval
liberally. On the other hand, the authors of Kaggle packages do not
use the feature at all. This point to a bifurcated user community:
experts who write package know how to use \eval while end-users do
not. An alternative explanation is that end-users do not need \eval
for simple data analysis tasks. This bifurcation was not observed in
the JavaScript community, but this may have changed with the advent of
Node.js ecosystem. The distribution of input sizes is similar to that
found in JavaScript, the majority of inputs being rather small.

\begin{figure}[!b]
	\centering
  \includegraphics[width=\columnwidth]{ast_sizes}
  \caption{Input sizes} \label{fig:ast-size}
\end{figure}

To roughly estimate the complexity of the code being evaluated,
Fig.~\ref{fig:ast-size} shows the distribution of the abstract syntax
tree size of the \c{expr} argument to \eval. The majority of inputs
across core and packages (\NbAstOnePercent of \eval calls and
\NbAstOneCallSitePercent of call sites) have size 1. That is to say,
the code being evaluated consists of a single variable name or
constant. A function call needs at least three nodes and a list such
as \c{c(1, 2, 3)} counts for 4 nodes. Inputs smaller than
\AstSizeNineFive nodes account for 95\% of all calls. Sizes in core
are larger than in packages. Most large ASTs originate from library
\c{VGAM}, which is used for statistical modeling, and also calls
functions using \eval from core. It uses \eval to evaluate large
matrices with values in a given environment.

\subsubsection{Consistency}

>>>>

\documentclass[USenglish,cleveref, autoref, thm-restate]{lipics-v2019}
\usepackage{listings,hyperref,multirow,paralist,xspace,url}
\input{macro}
\newcommand{\mypara}[1]{\medskip\noindent\emph{#1}\xspace}
\newcommand{\NOTE}[1]{{\it Note: #1}\xspace}

@@

\mypara{Eval in R.} The reflective interface exposed by
R is somewhat complex. The core library exports four functions with
slightly different semantics, \c{eval}, \c{evalq}, \c{eval.parent},
and \c{local}. \Eval is the more general function, and, unlike in
JavaScript, it takes three arguments, an expression, its environment
and the enclosing environment. The following discussion simplify some
details which are not relevant to this paper. For the interested
reader we recommend the excellent book by Hadley
Wickham~\cite{hadley}. The definition of \eval starts as follows:

@@

\mypara{CRAN Packages.} The packages that are included
in this study are the top \CorpusPackages packages based on their
reverse dependencies. The set of reverse dependencies for some package
\emph{P} is transitive closure of packages that import \emph{P}. The
hypothesis underlying that choice is that packages with a large set of
reverse depencies are likely to be higher in quality and their tests
will provide better coverage. The packages are implemented in R with
some C and Fortran code. Using \c{cloc}, we measure a total
\CorpusRCodeRnd lines of R code and \CorpusNativeCodeRnd lines of
native code. For each package, we use its runnable code (tests, etc.)
to compute code coverage. On average, the cde coverage is
\CorpusMeanExprCoverage which is acceptable without being exhaustive.
Figure~\ref{fig:corpus} shows these packages, the size of the dots
reflects the project's size in lines of code. The x-axis indicates
code coverage in percents and the y-axis gives the number of call
sites to \eval that were traced, in log scale. Dotted lines indicate
means. Packages with over \CorpusEvalsPackageTreshold eval call sites
are named.

@@

\mypara{Kaggle Scripts.} Kaggle is an online platform
for data-science and machine-learning. The website allows people to
submit data-analysis problems, users compete to find the best
solution. The solutions are uploaded to the platform as either plain
scripts or notebooks. We chose one of the most popular competition,
predicting the survival of passengers the
Titanic\footnote{\url{https://www.kaggle.com/c/titanic}}. Unlike CRAN,
Kaggle is not curated. After downloading the \CorpusKaggle solutions
and extracting the R code, we found that \CorpusDuplicatedKaggle were
duplicates. From the remaining \CorpusRunnableKaggle solutions,
\CorpusFailedKaggle failed to execute. Next to various runtime
exceptions, common problems were parse errors and misspelled package
names. The final set contains \CorpusFinishedKaggle programs
implemented in \CorpusFinishedKaggleCodeRnd lines of R code.

@@

\begin{figure}[!h]\hspace{-5mm}
  \includegraphics[width=1.05\linewidth]{pipeline.pdf}
  \caption{Analysis Pipeline}\label{fig:pipeline}
\end{figure}

@@

\mypara{Limitations.} Dynamic analysis can only observe
calls that are triggered by the program's input. We believe that
focusing on R packages with high code coverage does mitigate this to
some extent. The results we report here were obtained with R's
bytecode compiler turned off, this should not affect the results as
the compiler does not optimize \eval.

@@

@@

@@

@@

5(?) examples from real code.

@@

\subsection{Discussion}

Yeah, why?

\subsection{Can we do without?}

Similarly to JavaScript, there are also unnecessary uses of \eval. For
example, the \c{PerformanceAnalytics} package contains a function
\c{chart.QQPlot} that uses \eval to resolve a string into function and
another to call it and assign its results into a variable:
\begin{lstlisting}
function (R, d="norm", dp, ...) {
q.f <- eval(parse(text=paste("q",d,sep="")))
z <- NULL
eval(parse(text=paste("z<-q.f(",dp,",...)")))
}
\end{lstlisting}
  In both cases, there is no need for \eval:
\begin{lstlisting}
function (R, d="norm", dp, ...) {
q.f <- get(paste0("q",d))
z <- q.f(dp, ...)
}
\end{lstlisting}
  or even to a oneliner \c{do.call(paste0("q",d), as.list(dp, ...))}.

@@

Comparing the use of \eval in R and JavaScript is instructive. The use
of \eval is more widespread in data science code than it was in
JavaScript. Any run of an R program seems to trigger \eval; moreover
package programmers use \eval liberally. On the other hand, the
authors of Kaggle packages do not use the feature at all. This point
to a bifurcated user community: experts who write package know how to
use \eval while end-users do not. An alternative explanation is that
end-users do not need \eval for simple data analysis tasks. This
bifurcation was not observed in the JavaScript community, but this may
have changed with the advent of Node.js ecosystem. The distribution of
input sizes is similar to that found in JavaScript, the majority of
inputs being rather small.
